Multiplication, division and factorization\\of p-adic polynomials

### 2015-12-04 polynomials/muldivfact.tex
Proof of main algorithm
Getting closer to a first version of the proof.

\begin{rem}
Using $P \mapsto P^*=X^{deg(P)}P \left( \frac{1}{X} \right), $ the assumptions in \ref{theo:slope-factor} do not induce any loss in generality.
\end{rem}

@@

\begin{lem}
Some definitions :
\begin{align}
R_i &= A_{i+1} -A_i, \label{eqdef:Ri}\\
S_i &= P \% A_i, \label{eqdef:Si} \\
Q_i &= V_i P \div A_i. \label{eqdef:Qi}. 
\end{align}
Some formulae :
\begin{align}
B_i-B_{i+1} &= R_i B_{i+1} \div A_i, \label{eqdef:Biminus} \\
S_i-S_{i+1} &= R_i B_{i+1} \div A_i,  \label{eqdef:Siminus} \\
S_{i} &= (B_i R_i + (1-V_i B_i) S_{i-1}+(1-V_i B_i)(S_i-S_{i-1})) \% A_{i}, \label{eqdef:Si2}  \\
V_{i+1}-V_i &= (V_i (1-V_i B_i)+V_i^2(B_i-B_{i+1})) \% A_{i}, \label{eqdef:Viminus}\\
1-Q_i &= (1-V_i B_i)+((R_i-V_i S_i) \div A_i), \label{eqdef:Qiminus}\\
R_{i+1} &= ((V_{i+1}-V_i)S_{i+1}+(1-Q_i)R_i) \% A_{i+1}, \label{eq:Riplus}\\
1-V_{i+1}B_{i+1} &= ((1-V_i B_i)+V_i (B_i-B_{i+1}))^2. \label{eq:ViBiplus}
\end{align}
\end{lem}
\begin{proof}
We first remark that by definition,
\begin{align*}
P &= A_i B_i + S_i, \\
P V_i&= A_i Q_i + R_i.
\end{align*}
Since $P= A_i B_i + S_i= A_{i+1} B_{i+1} + S_{i+1},$ we have \[R_i B_{i+1}= (B_i -B_{i+1})A_i + (S_i+S_{i+1}).\]
Hence, by consideration of degree, what we have written is the Euclidean division of $R_i B_{i+1}$ by $A_i$ and we obtain \eqref{eqdef:Biminus} and \eqref{eqdef:Siminus}.

@@

\end{lem}

@@

\begin{lem}
$\NP (V_i)$ for $i>0$ ($V_0=1$):
\[\NP(V_i)=0+(d-1)\slp( \lambda_1). \]

@@

\draw (0,0) -- (1.5,-.375) -- (2,.625);
\node[below] at (.5,-0.25) {$\lambda_1$};
\node[below] at (2,-0.25) {$\lambda_2$};
\end{tikzpicture}
\end{lem}

@@

\begin{lem}
\begin{align*}
\NP (R_i), \NP(S_i) & \geq -\lambda_0+2^i(\lambda_1-\lambda_0)+(d-1) \slp (\lambda_0), \\
\NP (1-V_i B_i) & \geq 2^i (\lambda_1-\lambda_0)+(d-1) \slp (\lambda_0),\\
\NP (B_i-B_{i+1}) & \geq 2^i (\lambda_1-\lambda_0)+(\lambda_1-\lambda_0)+(l_1-1) \slp (\lambda_1)+\RS, \\
\NP (S_i-S_{i+1}) & \geq -\lambda_0 d+2^i (\lambda_1-\lambda_0)+(d-1) \slp (\lambda_0),\\
\NP (V_i-V_{i+1}) & \geq 2^i (\lambda_1-\lambda_0)+(d-1) \slp (\lambda_0),\\
\NP (1-Q_i) & \geq  2^i (\lambda_1-\lambda_0)+(d-1) \slp (\lambda_0).\\
\end{align*}
\end{lem}

>>>>

\begin{rem}
Using $P \mapsto P^*=X^{deg(P)}P \left( \frac{1}{X} \right), $ the assumptions in Theorem \ref{theo:slope-factor} do not induce any loss in generality.
\end{rem}

\begin{rem}
Non necessity to have $A_0$ as the first slope. It is possible to use minorations.

\draw (0,1)-- (1.5,0) -- (3,-.375)-- (3.5,.625);
\draw[green] (0,1.5)-- (1,0.33)-- (1.5,0) --(2,-.125)-- (3,0)-- (3.5,1);

\node[below] at (.75,-0.2) {$\lambda_0$};
\end{tikzpicture}
\end{rem}

@@

\begin{lem} \label{lem:formulae}
Some definitions :
\begin{align}
R_i &= A_{i+1} -A_i, \label{eqdef:Ri}\\
S_i &= P \% A_i, \label{eqdef:Si} \\
Q_i &= V_i P \div A_i. \label{eqdef:Qi}. 
\end{align}
Some formulae :
\begin{align}
B_i-B_{i+1} &= R_i B_{i+1} \div A_i, \label{eqdef:Biminus} \\
S_i-S_{i+1} &= R_i B_{i+1} \div A_i,  \label{eqdef:Siminus} \\
S_{i} &= (B_i R_i + (1-V_i B_i) S_{i-1}+(1-V_i B_i)(S_i-S_{i-1})) \% A_{i}, \label{eqdef:Si2}  \\
V_{i+1}-V_i &= (V_i (1-V_i B_i)+V_i^2(B_i-B_{i+1})) \% A_{i}, \label{eqdef:Viminus}\\
1-Q_i &= (1-V_i B_i)+((R_i-V_i S_i) \div A_i), \label{eqdef:Qiminus}\\
R_{i+1} &= ((V_{i+1}-V_i)S_{i+1}+(1-Q_i)R_i) \% A_{i+1}, \label{eq:Riplus}\\
1-V_{i+1}B_{i+1} &= ((1-V_i B_i)+V_i (B_i-B_{i+1}))^2. \label{eq:ViBiplus}
\end{align}
\end{lem}
\begin{proof}
We first remark that by definition,
\begin{align*}
P &= A_i B_i + S_i, \\
P V_i&= A_i Q_i + R_i.
\end{align*}
Since $P= A_i B_i + S_i= A_{i+1} B_{i+1} + S_{i+1},$ we have \[R_i B_{i+1}= (B_i -B_{i+1})A_i + (S_i+S_{i+1}).\]
Hence, by consideration of degree, what we have written is the Euclidean division of $R_i B_{i+1}$ by $A_i$ and we obtain \eqref{eqdef:Biminus} and \eqref{eqdef:Siminus}.

@@

\label{lem:initial_NP}
\end{lem}
\begin{proof}
Clear by Theorem \ref{theo:EDivisionNP}, and the fact that $coeff(P,d)=1.$
\end{proof}

@@

\begin{lem} \label{lem_unchanged_NP}
$\NP (V_i)$ for $i>0$ ($V_0=1$):
\[\NP(V_i)=0+(d-1)\slp( \lambda_1). \]

@@

\draw (0,0) -- (1.5,-.375) -- (2,.625);
\node[below] at (.5,-0.25) {$\lambda_1$};
\node[below] at (2,-0.25) {$\lambda_2$};
\end{tikzpicture}

$\NP (A_i)$:
\[\NP(A_i)=-d \lambda_0+d \slp( \lambda_0). \]

\draw (0,1.5)-- (1.5,0) ;
\node[below] at (.75,0.5) {$\lambda_0$};
\end{tikzpicture}
\end{lem}

@@

\begin{lem} \label{lem:iteration_inside_the_loop}
\begin{align*}
\NP (R_i), \NP(S_i) & \geq -\lambda_0+2^i(\lambda_1-\lambda_0)+(d-1) \slp (\lambda_0), \\
\NP (1-V_i B_i) & \geq 2^i (\lambda_1-\lambda_0)+(d-1) \slp (\lambda_0),\\
\end{align*}
\end{lem}

We prove the last two lemmas by induction.

\begin{proof}
The case $i=0$ is clear thanks to Theorem \ref{theo:EDivisionNP} and Lemma \ref{lem:initial_NP}.

Now, we assume that for some $i \geq 0$ the formulae defined in the Lemmae \ref{lem_unchanged_NP} and \ref{lem:iteration_inside_the_loop} are satisfied. 
We prove them for $i+1$.

Firstly, thanks to the minoration on $\NP (R_i)$ and the fact that $R_i=A_{i+1}-A_i$, we have $\NP (A_{i+1}= \NP (A_i).$
Since $B_{i+1} = P \div A_{i+1},$ we have $\NP (B_{i+1}) = \NP (B_i).$
We apply \eqref{eqdef:Biminus} to obtain 
\[\NP (B_i-B_{i+1})  \geq 2^i (\lambda_1-\lambda_0)+(\lambda_1-\lambda_0)+(l_1-1) \slp (\lambda_1)+\RS. \]
Then, we can apply \eqref{eqdef:Siminus} to obtain 
\[\NP (S_i-S_{i+1})  \geq -\lambda_0 d+2^i (\lambda_1-\lambda_0)+(d-1) \slp (\lambda_0).\]

As a consequence, we get that $\NP (S_{i+1}) \geq \NP (S_i)$ (pointwise).

Now, we can apply \eqref{eqdef:Viminus} to obtain that 
\[\NP (V_i-V_{i+1}) \geq 2^i (\lambda_1-\lambda_0)+(d-1) \slp (\lambda_0).\]
As a consequence, we get that $\NP(V_i)=\NP(V_{i+1}).$
Similarly, we use \eqref{eqdef:Qiminus} to obtain that
\[\NP (1-Q_i)  \geq  2^i (\lambda_1-\lambda_0)+(d-1) \slp (\lambda_0).\]

This is enough to conclude on the Newton polygon of $R_{i+1}.$
Indeed, thanks to \eqref{eq:Riplus}, we obtain that 
\[\NP (R_i) \geq -\lambda_0+2^i(\lambda_1-\lambda_0)+(d-1) \slp (\lambda_0). \]
We can proceed similarly for $1-V_{i+1} B_{i+1}$ using \eqref{eq:ViBiplus}.

Finally, we can apply \eqref{eqdef:Si2} for $i+1,$
\[S_{i+1} = (B_{i+1} R_{i+1} + (1-V_{i+1} B_{i+1}) S_{i}+(1-V_{i+1} B_{i+1})(S_{i+1}-S_{i})) \% A_{i+1},\] to obtain the desired minoration on $\NP (S_{i+1}).$ This concludes the proof.
\end{proof}
